---
title: "ARM Exercise 2"
output: html_document
date: "2025-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r library}
# import libraries here
library(viridis)
library(arules)
library(TSP)
library(data.table)
#library(ggplot2)
#library(Matrix)
library(tcltk)
library(dplyr)
library(devtools)
library(purrr)
library(tidyr)
library(arulesViz)
```


```{r}
# Import raw tokenized data, partially cleaned in excel
data <- read.csv('TokenizedData.csv', stringsAsFactors = FALSE)
head(data$Tokens)
head(data$LABEL)
```

Hmmm, there's a space after every row of text data. That needs to be removed:

```{r}
# Use trim white space, and gsub to replace any unusual characters
data$Tokens <- trimws(data$Tokens)
data$Tokens <- gsub("\\s+", " ", data$Tokens)
```


```{r}
# transform data into list for via strsplit
transaction_list <- strsplit(data$Tokens, " ", fixed = TRUE)

# Convert to transactions
transactions <- as(transaction_list, "transactions")
```
Next, the data will be cleaned to remove words smaller than 4 characters (often stopwords)

```{r}
# Remove words less than 4 characters, via lapply
trans_list_fill <- lapply(transaction_list, function(x) x[nchar(x) >= 4])

# revert data back to transactions
transactions2 <- as(trans_list_fill, 'transactions')

```





```{r}
# check results
inspect(head(transactions2))
```
Good! The data is in transaction basket format. This was accomplished by transforming the token lists into lists again
(to work with R formatting), and then cleaning them, and turning them back into baskets with as.transactions.

Now that it's in transaction style, the data can be run through arules apriori. I decided on a confidence of 0.5 
and a support of 0.35, which I found to have few enough rules to be meaningful to the dataset.

```{r}
# Now that data is clean/in transaction format, apply apriori with support 0.035 and confidence 0.5
Frules = apriori(transactions2, parameter = list(support=.035, 
                                                 confidence=.5, minlen=2))
```

Next, I'll look at the top 15 rules by confidence, support, and lift:

```{r}
# rules by confidence
SortedRulesC <- sort(Frules, by="confidence", decreasing=TRUE)
inspect(SortedRulesC[1:15])

```

Interesting! Healthcare is always the consequent. This make sense, since it's a database of sentiment on universal healthcare.

Now support:

```{r}
# Rules by support
SortedRulesS <- sort(Frules, by="confidence", decreasing = TRUE)
inspect(SortedRulesS[1:15])
```

All the rules are also healthcare-dependent. The high prevalence of support for rules with 2-item antecedents is neat.
They're words typically associated with healthcare and the U.S. healthcare system. I think rule 4 is most interesting because of 'don't'.

```{r}
# Rules by lift
SortedRulesL <- sort(Frules, by = 'lift', decreasing = TRUE)
inspect(SortedRulesL[1:15])
```

The only non-healthcare consequents! I find it interesting that one rule is an antecedent of 'want' and consequent of 'don't'. 

```{r}
# check overall rule results
inspect(Frules)

```


The total set of rules for this set is 39. Not a ton. The basic unordered version has more of a focus on money/countries too.

Time for plots!

```{r}
# Time to make visualizations

# Plot of which items are most frequent

itemFrequencyPlot(transactions2, topN = 20,
                          col = brewer.pal(8, 'Pastel2'),
                          main = 'Relative Item Frequency Plot',
                          type = "relative", 
                          ylab = "Item Frequency (Relative)")

# Sort rules by lift
SortedRules <- sort(Frules, by="lift", decreasing=TRUE)
inspect(SortedRules[1:10])
(summary(SortedRules))

```

Cost comes up a bit here, main focus is also on healthcare/insurance as ideas. Also seems to be a lot of 'opinion-based' words here. Cool.

```{r}
# Visualize
# tcltk lift

subrules <- head(sort(SortedRules, by="lift"),10)
plot(subrules)

plot(subrules, method="graph", engine="interactive")
```

```{r}
# Visualize
# tcltk support


subrules <- head(sort(SortedRules, by="support"),10)
plot(subrules)

#plot(subrules, method="graph", engine="interactive")
plot(subrules, method="graph", engine="htmlwidget")
```

```{r}
# Visualize
# tcltk confidence


subrules <- head(sort(SortedRules, by="confidence"),10)
plot(subrules)

#plot(subrules, method="graph", engine="interactive")
plot(subrules, method="graph", engine="htmlwidget")
```





