---
title: "R Healthcare HClust"
output: html_document
date: "2025-03-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r library}
library(tm)
library(stringr)
library(wordcloud)

library(slam)
library(quanteda)

library(SnowballC)
library(arules)

library(proxy)
library(cluster)
library(stringi)
library(proxy)
library(Matrix)
library(tidytext) # convert DTM to DF
library(plyr) ## for adply
library(ggplot2)
library(factoextra) # for fviz
library(mclust) # for Mclust EM clustering
library(textstem)  ## Needed for lemmatize_strings

library(amap) 
library(networkD3)
```


```{r}
# load in csv file of raw text content
data <- read.csv('Raw Master Data.csv', stringsAsFactors = FALSE)
# check
head(data)
```

```{r}
head(data$Content)
```

```{r}
# Turn content column into corpus with tm library
ContentCorpus <- Corpus(VectorSource(data$Content))

(getTransformations()) 
(ndocs<-length(ContentCorpus))

ContentCorpus <- tm_map(ContentCorpus, content_transformer(tolower))
ContentCorpus <- tm_map(ContentCorpus, removePunctuation)
## Remove all Stop Words
ContentNovelsCorpus <- tm_map(ContentCorpus, removeWords, stopwords("english"))

# Taking out common words
MyStopWords <- c("and","like", "very", "can", "I", "also", "lot", "may", "us") # US is a common one, but not useful due to popularity
ContentCorpus <- tm_map(ContentCorpus, removeWords, MyStopWords)
ContentCorpus <- tm_map(ContentCorpus, lemmatize_strings)

# check results
summary(ContentCorpus)
```

```{r}
Content_dtm <- DocumentTermMatrix(ContentCorpus,
                         control = list(
                           stopwords = TRUE, ## remove normal stopwords
                           wordLengths=c(4, 15),
                           removePunctuation = TRUE,
                           removeNumbers = TRUE,
                           tolower=TRUE,
                           #stemming = TRUE,
                           remove_separators = TRUE))

DTM_mat <- as.matrix(Content_dtm)
DTM_mat[1:13,1:10]
```

```{r}
# Check word frequency
(WordFreq <- colSums(as.matrix(Content_dtm)))

(head(WordFreq))
(length(WordFreq))
ord <- order(WordFreq)
(WordFreq[head(ord)])
(WordFreq[tail(ord)])
## Row Sums
(Row_Sum_Per_doc <- rowSums((as.matrix(Content_dtm))))
```


```{r}
# Copy the matrix
Content_M <- as.matrix(Content_dtm)
(Content_M[1:13,1:5])

# Normalize the matrix
Content_M_N1 <- apply(Content_M, 1, function(i) round(i/sum(i),3))
(Content_M_N1[1:13,1:5])

# Transpose data
Content_Matrix_Norm <- t(Content_M_N1)
(Content_Matrix_Norm[1:13,1:10])
```
```{r}
as.matrix(Content_dtm[1:10, 1:10])
Content_M <- as.matrix(Content_dtm)
Content_DF <- as.data.frame(Content_M)
```

```{r}
Content_DF[1:3, 1:10]
(nrow(Content_DF))  # check number of news headlines/reddit pulls
# DF formatting
ncol(Content_DF)
# Convert normalized data
Content_DF_From_Matrix_N<-as.data.frame(Content_Matrix_Norm)
```

```{r}
# check data first
(Content_Matrix_Norm[c(1:6),c(3850:3900)])

# also check non-normalized data to compare
(Content_M[c(1:6),c(3850:3900)])
```

```{r}
# make data copies
m  <- Content_M
m_norm <-Content_Matrix_Norm
(str(m_norm))
```

```{r}
distMatrix_C <- dist(m, method="cosine")
print("cos sim matrix is :\n")
print(distMatrix_C[1:10]) #check results

# Normalized data version
distMatrix_C_norm <- dist(m_norm, method="cosine", upper = TRUE, diag = TRUE)
print("The norm cos sim matrix is :\n")
print(distMatrix_C_norm[1:10])
```

```{r}
# clustering
groups_C <- hclust(distMatrix_C,method="ward.D")


plot(groups_C, cex=.7, hang=-30,main = "Cosine Sim")
rect.hclust(groups_C, k=5)

radialNetwork(as.radialNetwork(groups_C))
dendroNetwork(groups_C)
```
`
Hmmm. It's hard to read. I'll try recoloring:

```{r}
# recolor using dendextend library
avg_dend_obj <- as.dendrogram(groups_C)
avg_col_dend <- color_branches(avg_dend_obj, h = 5)
plot(avg_col_dend, main = 'Full Dendrogram')
```

```{R}
# Normalized clustering
# Clustering the full dataset 

groups_C_n <- hclust(distMatrix_C_norm,method="ward.D")
plot(groups_C_n, cex=0.9, hang=-1,main = "Cosine Sim and Normalized")
rect.hclust(groups_C_n, k=5)

radialNetwork(as.radialNetwork(groups_C_n))
```

```{r}
# Going to try a randomized sample of 200 and re-run the dendrogram
# sample without replacement
data_sample <- data[sample(1:nrow(data), 50, replace = FALSE),]
```

Next, re-run calcuations on the sampled data:

```{r}
# Turn content column into corpus with tm library
ContentCorpus2 <- Corpus(VectorSource(data_sample$Content))

(getTransformations()) 
(ndocs<-length(ContentCorpus2))

ContentCorpus2 <- tm_map(ContentCorpus2, content_transformer(tolower))
ContentCorpus2 <- tm_map(ContentCorpus2, removePunctuation)
## Remove all Stop Words
ContentNovelsCorpus2 <- tm_map(ContentCorpus2, removeWords, stopwords("english"))

# Taking out common words
MyStopWords <- c("and","like", "very", "can", "I", "also", "lot", "may", "us") # US is a common one, but not useful due to popularity
ContentCorpus2 <- tm_map(ContentCorpus2, removeWords, MyStopWords)
ContentCorpus2 <- tm_map(ContentCorpus2, lemmatize_strings)

# check results
summary(ContentCorpus2)
```

```{r}
Content_dtm2 <- DocumentTermMatrix(ContentCorpus2,
                         control = list(
                           stopwords = TRUE, # remove normal stopwords
                           wordLengths=c(4, 15), # remove words less than 4 or greater than 15 characters
                           removePunctuation = TRUE,
                           removeNumbers = TRUE,
                           tolower=TRUE,
                           #stemming = TRUE,
                           remove_separators = TRUE))

DTM_mat2 <- as.matrix(Content_dtm2)
DTM_mat2[1:13,1:10]
```
```{r}
# Check word frequency
(WordFreq2 <- colSums(as.matrix(Content_dtm2)))

(head(WordFreq2))
(length(WordFreq2))
ord <- order(WordFreq2)
(WordFreq2[head(ord)])
(WordFreq2[tail(ord)])
## Row Sums
(Row_Sum_Per_doc2 <- rowSums((as.matrix(Content_dtm2))))
```

```{r}
# Copy the matrix
Content_M2 <- as.matrix(Content_dtm2)
(Content_M2[1:13,1:5])

# Normalize the matrix
Content_M_N12 <- apply(Content_M2, 1, function(i) round(i/sum(i),3))
(Content_M_N12[1:13,1:5])

# Transpose data
Content_Matrix_Norm2 <- t(Content_M_N12)
(Content_Matrix_Norm2[1:13,1:10])
```
```{r}
as.matrix(Content_dtm2[1:10, 1:10])
Content_M2 <- as.matrix(Content_dtm2)
Content_DF2 <- as.data.frame(Content_M2)
```

```{R}

Content_DF2[1:3, 1:10]
(nrow(Content_DF2))  # check number of news headlines/reddit pulls
# DF formatting
ncol(Content_DF2)
# Convert normalized data
Content_DF_From_Matrix_N2<-as.data.frame(Content_Matrix_Norm2)
```

```{R}
# check data first
(Content_Matrix_Norm2[c(1:6),c(140:190)])

# also check non-normalized data to compare
(Content_M2[c(1:6),c(140:190)])
```

```{r}
# make data copies
m2  <- Content_M2
m_norm2 <-Content_Matrix_Norm2
(str(m_norm2))
```

```{r}
# Make new distance matrices
distMatrix_C2 <- dist(m2, method="cosine")
print("sample cos sim matrix is :\n")
print(distMatrix_C2[1:10]) #check results

# Normalized data version
distMatrix_C_norm2 <- dist(m_norm2, method="cosine", upper = TRUE, diag = TRUE)
print("The sample norm cos sim matrix is :\n")
print(distMatrix_C_norm2[1:10])
```
```{r}
# re-do clustering with sampled data version
groups_C2 <- hclust(distMatrix_C2,method="ward.D")


plot(groups_C2, cex=.7, hang=-30,main = "Cosine Sim")
rect.hclust(groups_C2, k=5)

radialNetwork(as.radialNetwork(groups_C2))
dendroNetwork(groups_C2)
```

Neat! It's much eaiser to see the results on smaller scales with dendrograms.
The divisions seem to suggest that most of the inforamtion is located in 3 distinct clusters,
So k = 3 would be an optimal value to try.

Next, I'll recolor this version:

```{r}
# recolor using dendextend library
avg_dend_obj2 <- as.dendrogram(groups_C2)
avg_col_dend2 <- color_branches(avg_dend_obj2, h = 3)
plot(avg_col_dend2, main = 'Sampled Data Dendrogram')
```

Hmmm. No clue why that works weird. Time to try the reduced normalized data.

```{r}
# Normalized reduced data clustering
groups_C_n2 <- hclust(distMatrix_C_norm2,method="ward.D")
plot(groups_C_n2, cex=0.9, hang=-1,main = "Reduced Cosine Sim and Normalized")
rect.hclust(groups_C_n2, k=3)

radialNetwork(as.radialNetwork(groups_C_n2))

```





















